---
title: "Homework 4"
author: "Caitlin Cassidy"
date: "10/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1:

Build a glm in R to classify individuals as either Male or Female based on their weight and height. What is the accuracy of the model?


```{r q1}
#read in csv
index <- read.csv("hw_data/500_Person_Gender_Height_Weight_Index.csv", header = TRUE)

#function to split data into train test validate
model_split <- function(dfi, train_p, validate_p, test_p, col_name="exp_group"){
    dfi <- sample_n(dfi, nrow(dfi),replace=FALSE);
    p <- (seq(nrow(dfi))-1)/nrow(dfi);
    train_dfi <- dfi %>% filter(p < train_p);
    validate_dfi <- dfi %>% filter(p < train_p + validate_p & p >= train_p);
    test_dfi <- dfi %>% filter(p >= train_p + validate_p);
    train_dfi[[col_name]] <- "train";
    validate_dfi[[col_name]] <- "validate";
    test_dfi[[col_name]] <- "test";
    rbind(train_dfi, validate_dfi, test_dfi);
}

#function to get f1 score
f1 <- function (y_true, y_pred, positive = NULL) 
{
    Confusion_DF <- ConfusionDF(y_pred, y_true)
    if (is.null(positive) == TRUE) 
        positive <- as.character(Confusion_DF[1, 1])
    Precision <- Precision(y_true, y_pred, positive)
    Recall <- Recall(y_true, y_pred, positive)
    F1_Score <- 2 * (Precision * Recall)/(Precision + Recall)
    return(F1_Score)
}

#create indicator variable for gender
index <- index %>% mutate(gender1 = ifelse(Gender == "Male", 0, 1))

#split data into train test validate
index <- rbind(model_split(index, 1/3, 1/3, 1/3))
index %>% group_by(exp_group) %>% tally()

#split data into 3 groups
train <- index %>% filter(exp_group=="train")
validate <- index %>% filter(exp_group=="validate")
test <- index %>% filter(exp_group=="test")

#create logistic model on train data
model <- glm(gender1 ~ Height + Weight, family=binomial(link='logit'), data=train)

#
pred <- predict(model, newdata=validate, type="response")
pred
sum((pred>0.5) == validate$gender1)/nrow(validate)

#f1 score
f1(validate$gender1, pred > 0.5)
```

This model has an accuracy of about 53.3%.

## Problem 2:

Use the ‘gbm’ package to train a similar model. Don’t worry about hyper parameter tuning for now.
What is the accuracy of the model?

```{r q2}
#install.packages("gbm")
library(gbm)

model <- gbm(gender1 ~ Height + Weight, distribution="bernoulli",
             data=train,
             n.trees = 100,
             interaction.depth = 2,
             shrinkage = 0.1)

pred <- predict(model, validate, type="response")
sum((pred>0.5)==validate$gender1)/nrow(validate)

```
The accuracy of this model is about 53.9%

## Problem 3:

Filter the data set so that it contains only 50 Male examples. Create a new model for this data set. What is the F1 Score of the model?

## Problem 4:

For the model in the previous example plot an ROC curve. What does this ROC curve mean?


## Problem 5:

Using K-Means, cluster the same data set. Can you identify the clusters with the known labels? Provide an interpretation of this result.
