---
title: "Biodiversity in National Parks"
author: "Caitlin Cassidy"
date: "10/1/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction to this Project
The United States has 62 National Parks. Data on 56 of these parks was provided by the National Park Service. Because of the varying geography of the U.S., the parks vary greatly in terms of their biodiversity. Biodiversity refers to the variety and variability of species in an ecosystem. Biodiversity is important for maintaining healthy ecosystems.

In this project, we will ????

## Number of Unique Species by Latitude 
The U.S. National Parks are home to a variety of different species of mammals, birds, reptiles, fish, plants and other types of organisms. Since these parks are located all over the U.S. the parks will vary in terms of what species and how many they are home to. 

If we consider the number of unique species for different ranges of latitude, we can see how many species live in each range of latitude.
```{r latitude plot}
#plot(lat_figure)
```
By looking at these bar charts, we can see that there are greater numbers of unique mammal, bird and reptile species that live in latitudes in the middle latitudes. Very few mammal, bird and reptile species live in low latitudes or high latitudes. The climate conditions at the low and high latitudes are more extreme than climates in the middle latitudes. Therefore, this data suggests that middle latitudes provide a better climate for mammals, birds and reptiles to live.
The number of unique fish species, however, does not follow this same pattern. While more fish species live in low latitudes than middle or high latitudes, the distribution is more even across all latitudes than the distribution of the other species.



#Figures to Make
correllation matrix
scatterplots
number of species by lattitude

correlation between prey/predators
most common species in a few parks
glm for predicting something based on number of species
conservation
