---
title: "Biodiversity in National Parks"
author: "Caitlin Cassidy"
date: "10/1/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction to this Project
The United States has 62 National Parks. Data on 56 of these parks have been provided by the National Park Service. This data details biodiversity within the parks. The data contain a catalog of species that live within the parks, information on the conservation status of many of these species, as well as general info regarding the parks' locations and size.

Because of the varying geography of the U.S., the parks vary greatly in terms of their biodiversity. Biodiversity refers to the variety and variability of species in an ecosystem. Biodiversity is important for maintaining healthy ecosystems. It is important that we keep records of park biodiversity in order to understand the ecology and health of these ecosystems.

In this project, we will examine this data set in an effort to uncover patterns and relationships in the data. We will also generate a few linear models and suggest efforts for future data collection in the National Park system.

## Variability of Number of Species Latitude 
The U.S. National Parks are home to a variety of different species of mammals, birds, reptiles, and fish, among many other types of organisms. Since these parks are located all over the U.S., they will vary in terms their wildlife. 

One interesting pattern that emerges in the data is the number of unique species at different latitudes. The following bar chart depict the number of unique species of mammals (A), birds (B), reptiles (C), and fish (D), at different latitudes. The latitude categories are defined in intervals of 10 degrees latitude. 

```{r latitude_plot, fig.width=10, fig.height=10}
readRDS("figures/lat_figure.rds")
```
By looking at these bar charts, we can see that there are greater numbers of unique mammal, bird and reptile species that live in latitudes in the middle latitudes. Very few mammal, bird and reptile species live in low latitudes or high latitudes. The climate conditions at the low and high latitudes are more extreme than climates in the middle latitudes. Therefore, this data suggests that middle latitudes provide a better climate for mammals, birds and reptiles to live. This pattern makes sense. It is also interesting to note that the distributions for mammals and birds are extremely similar. I will follow up on this relationship later in this discussion.

The number of unique fish species, however, does not follow this same pattern. While more fish species live in low latitudes than middle or high latitudes, the distribution is more even across the categories of latitude than the distribution of the other species.

In general, from these figures we can see that parks at middle latitudes have a great number of unique species and thus greater biodiversity than parks at low and high latitudes.

## Relationships Between Species
In many ecosystems, it has been observed that the number of bird species and mammal species are positively correlated (cite). I decided to see if a similar pattern emerged in the National Parks data. 

```{r mammals_scatterplots, echo = FALSE, warning = FALSE, message = FALSE}
readRDS("figures/scatter1.rds")
readRDS("figures/scatter2.rds")
```
These scatterplots demonstrate correlations between mammals and other species in the parks.The different size of the points represents the acreage of the park (larger points correspond to larger acreage). 

The first scatterplot shows a positive association between the number of mammal species and the number of bird species in the parks. 

The second scatterplot shows a negative association between the number of mammal species and the number of fish species in the parks.

From these plots, we can see that there are some associations between the number of mammal species and the number of other types of species. This information is useful for a variety of reasons. For example, it may be possible to estimate the number of mammal species in a park for which the number of bird and fish species are known.

## Can we predict the number of mammal species in a park?
Is it possible to predict the number of mammal species in a park with reasonable accuracy?
To answer this question, I used variable selection techniques to __. 


## Conservation Status of Species

```{r dot_plots, echo = FALSE, warning = FALSE, message = FALSE, fig.height=10}
readRDS("figures/dot_plot_e.rds")
```

```{r dot_plots_2, echo = FALSE, warning = FALSE, message = FALSE, fig.height=10}
readRDS("figures/dot_plot_c.rds")
```


## Future Work
The Biodiveristy in National Parks data set contains many pieces of useful information. I think that it will be beneficial to continue and expand data collection efforts within the National Park system. The data sets that were used for this project contained information on the number of unique species in the parks, but did not contain information about the number of unique organisms in the parks. I think that efforts should be undertaken to sample each of the National Parks to collect data on the individual organisms. This type of data collection would be a massive undertaking and would likely need to be a long term effort. I think it would also be very useful to track changes in populations over time. This would allow park rangers and the public to be aware of changing animal and plant populations. As more and more species are becoming endangered each year, tracking the number of individual organisms over time could allow us to determine if conservation efforts are successful or to predict how many years are left until another species is extinct.


